Study of the numerical solution of X-Modes equations around the hybrid resonnance

### 2015-01-27 report/numerics.tex
numbering in numerics.tex

The second part of the section deals with the question of the equivalence of the limiting absorption and limited amplitude 
principle. We compare the solutions obtained as $\nu\rightarrow 0$ with the help of our frequency domain code 
and of the time domain code (computed for large values of time). 
The time-domain code implements the scheme described in Section \ref{sec:discr} and 
is written in Fortran. 
\subsection{Frequency Domain Problem}
\label{sec:freq_dep}
\subsubsection{Validity of Implementation}
To check the validity of the code, we perform a numerical experiment with (formally chosen) parameters:
\begin{align}
\label{eq:parameters}
\alpha(x)=x^2+1,\qquad \delta(x)=\left(\alpha^2+x\alpha\right)^{\frac{1}{2}}.
\end{align}
Additionally, the boundary conditions read as 
\begin{align}
\label{eq:bcs}
\partial_{1}E_{y}(-L)+2iE_{y}(-L)=2iAi(-L)+Ai'(-L),\\
\partial_{1}E_{y}(H)=0,
\end{align}
where $Ai(x)$ is the Airy function. 
For more detail on Airy functions and the Airy equation see \cite[Chapter 10.4]{abramowitz_stegun}.

It can be shown that $E_{y}$ thus satisfies the Airy equation, and hence, 
with the choice of the boundary conditions as above, we obtain that 
\begin{align*}
E_{y}=Ai(x),\qquad 
E_{x}=-i\frac{\delta(x)}{\alpha(x)}Ai(x)
\end{align*}
is the solution to the problem with parameters (\ref{eq:parameters}) with the boundary condition (\ref{eq:bcs}).

In Figure \ref{fig:conv_rate} we demonstrate the convergence rates for this problem, with 
\begin{align*}
\|E_{x}-E_{x}^{c}\|_{L_{2}(\Omega)}=\left(\int\limits_{\Omega}\left|E_{x}(x)-\sum\limits_{k=1}^{N}e_{xk}\phi_{k}\right|^2dx\right),
\end{align*}
where $\left(\phi_{i}\right)_{i=1}^{N}$ are $P_{1}$-finite elements and $E_{x}(x)$ is a known analytic solution. Similarly $\|E_{y}-E_{y}^{c}\|_{L_{2}(\Omega)}$ and $\|E_{y}-E_{y}^{c}\|_{H_{1}(\Omega)}$ are defined.

@@

We use the following well-known facts:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq \nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness}.  
\item The form of the exact solution to the problem (\ref{vf1dcase}):
\begin{align}
\label{eq:exact}
 E_x^{\nu}=E_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f(x)}{\alpha(x)+i\nu},
\end{align}
for some $f(x)\in L_{2}(\Omega)$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:bsc}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2|v''|_{L_{2}(\Omega)},\; C>0,
 %\\ \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$.
\end{itemize}
For $E_x^{\nu},\;f(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\begin{align*}
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{f''}{\alpha+i\nu}-2\frac{f'\alpha'}{(\alpha+i\nu)^2}+\frac{f\alpha''}{(\alpha+i\nu)^3},
\end{align*}
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\begin{align*}
 \left|  \frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
 %\\
 %&=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\end{align*}
where $C>0$ does not depend on $\nu$. This, together with the estimates (\ref{eq:cea}) and (\ref{eq:bsc}), results in 
\begin{align*}
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2,
\end{align*}
from which it follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$.

@@

\begin{remark}
 We can show that 
\begin{align*}
 \|E^{\nu}_{x}\|_{L_{2}(\Omega)}\leq \frac{C}{\sqrt{\nu}},\; C>0, 
\end{align*}
and thus the relative error control
\begin{align*}
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}}{\|E^{\nu}_{x}\|_{L_{2}(\Omega)}}\leq \epsilon
\end{align*}
can be ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{4}}$.
\end{remark}

>>>>

The second part of the section deals with the question of the equivalence of the limiting absorption and limited amplitude 
principle. We compare the solutions obtained as $\nu\rightarrow 0$ with the help of our frequency domain code 
and of the time domain code (computed for large values of time). 
The time-domain code implements the scheme described in Section \ref{sec:discr} and 
is written in Fortran. 
\subsection{Frequency Domain Problem}
\label{sec:freq_dep}
\subsubsection{Validity of Implementation}
To check the validity of the code, we perform a numerical experiment with (formally chosen) parameters:
\begin{align}
\label{eq:parameters}
\alpha(x)=x^2+1,\qquad \delta(x)=\left(\alpha^2+x\alpha\right)^{\frac{1}{2}}.
\end{align}
Additionally, the boundary conditions read as 
\bealn
\label{eq:bcs}
\partial_{1}E_{y}(-L)+2iE_{y}(-L)=2iAi(-L)+Ai'(-L),\\
\partial_{1}E_{y}(H)=0,
\eealn
where $Ai(x)$ is the Airy function. 
For more detail on Airy functions and the Airy equation see \cite[Chapter 10.4]{abramowitz_stegun}.

It can be shown that $E_{y}$ thus satisfies the Airy equation, and hence, 
with the choice of the boundary conditions as above, we obtain that 
\bealn
E_{y}=Ai(x),\qquad 
E_{x}=-i\frac{\delta(x)}{\alpha(x)}Ai(x)
\ealn
is the solution to the problem with parameters (\ref{eq:parameters}) with the boundary condition (\ref{eq:bcs}).

In Figure \ref{fig:conv_rate} we demonstrate the convergence rates for this problem, with 
\ben
\|E_{x}-E_{x}^{c}\|_{L_{2}(\Omega)}=\left(\int\limits_{\Omega}\left|E_{x}(x)-\sum\limits_{k=1}^{N}e_{xk}\phi_{k}\right|^2dx\right),
\een
where $\left(\phi_{i}\right)_{i=1}^{N}$ are $P_{1}$-finite elements and $E_{x}(x)$ is a known analytic solution. Similarly $\|E_{y}-E_{y}^{c}\|_{L_{2}(\Omega)}$ and $\|E_{y}-E_{y}^{c}\|_{H_{1}(\Omega)}$ are defined.

@@

We use the following well-known facts:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq \nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness}.  
\item The form of the exact solution to the problem (\ref{vf1dcase}):
\begin{align}
\label{eq:exact}
 E_x^{\nu}=E_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f(x)}{\alpha(x)+i\nu},
\end{align}
for some $f(x)\in L_{2}(\Omega)$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:bsc}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2|v''|_{L_{2}(\Omega)},\; C>0,
 %\\ \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$.
\end{itemize}
For $E_x^{\nu},\;f(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{f''}{\alpha+i\nu}-2\frac{f'\alpha'}{(\alpha+i\nu)^2}+\frac{f\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\ben
 \left|  \frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
 %\\
 %&=c\left.
 %\left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\een
where $C>0$ does not depend on $\nu$. This, together with the estimates (\ref{eq:cea}) and (\ref{eq:bsc}), results in 
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2,
\een
from which it follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$.

@@

\begin{remark}
 We can show that 
\bealn
 \|E^{\nu}_{x}\|_{L_{2}(\Omega)}\leq \frac{C}{\sqrt{\nu}},\; C>0, 
\eealn
and thus the relative error control
\bealn
 \frac{\|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}}{\|E^{\nu}_{x}\|_{L_{2}(\Omega)}}\leq \epsilon
\eealn
can be ensured by choosing $h$ as $\beta_{\epsilon}\nu^{\frac{3}{4}}$.
\end{remark}
